---
title: "r4ds Solutions"
author: "Dan Ovando"
date: "August 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)
library(purrr)
library(stringr)


```

# 3.35 - faceting

```{r 3.35 faceting}

mpg %>% 
  ggplot(aes(year,cty)) + 
  geom_point() + 
  facet_wrap(~hwy)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = manufacturer, shape =  trans), show.legend = F) +
  facet_grid(. ~ cyl)

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), position = "jitter")

ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_count()
  
geom_point(position = 'jitter')

```

Interesting, the `..X..` stuff tells ggplot that its refering to a `stat_` transformation of the data, instead of the input

There are three positions: `identity` (take the data literally, think bar plots), `fill` which essentially stacks things, `dodge` which places things next to each other, and `jitter` which moves things around. Manipulate outisde of aes per `geom_bar(aes(), position = dodge)`

geom_point counts number of observations at a given location, for when you've got overlapping things

`coord_quickmap` moves a polygon into the right projection for maps

```{r}

nz <- map_data("nz")


ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")

ggplot(nz, aes(long, lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_quickmap()

```


# 3.9.1

```{r}

mpg %>% 
  ggplot(aes(manufacturer, fill = drv)) + 
  geom_bar()


mpg %>% 
  ggplot(aes(manufacturer, fill = drv)) + 
  geom_bar() + 
  coord_polar() + 
  labs(x = 'hello', title = 'weak')


ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
 
```

Interesting, using `coord_fixed` holds the aspect ratio, so that its 1:1 (or whatever you want), so that for example an identity line is at a 45 degree angle. 

# 8.2.2

`\n` in a text tile specifies newline, so the thing below has top row x,y, next row 1,'a,b'


```{r}
read_delim("x,y\n1,'a,b'", quote = "'", delim = ',')

```

# 10.3

```{r}
library(nycflights13)

planes %>% 
  count(tailnum) 

```
 
 
# 10.3
 
```{r regex}
library(readr)
 
 x <- c("apple", "banana", "pear")
str_view(x, "an")
 
str_view(x, '.a.')


 
```

OK! `.` tells it to match anything. so `.a.` says finds the two characters with an `a` in the middle,  `..a` finds anything with two characters then an a

`\` is an escape, basically saying don't do something. So, `\.` means take `.` literally, instead of using it as a command. But `\` is a command itself, so `\\` tells R not to use it as a command

Cool! So `^` says beginning, `$` says end, `.` says anything, `.*` says everything before/after thing

```{r}
str_view(c("abc", "a.c", "bef"), "a\\.c")


x <- "a\\b"

str_view(x,"\\\\")

str_view(c("a\b"), "\\\\")
```

# 11.3.2.1

```{r}

# Match literal string "$^$"

b = "ab$^$b"

str_match(b, "\\$.*")

str_match(b, ".*\\$")

# 2 starts with y or ends with x
 
a = stringr::words

str_subset(a, pattern = '^y')

str_subset(a, pattern = 'y$')

str_length(a)

# Is exactly 3 things long
str_view(a, pattern = '^...$', match = T)

str_subset(a, pattern = '^...$')

# Is 7 or more chacaters long

str_view(a, pattern = '^.......', match = T)

str_subset(a, pattern = '^.......')


```

# 11.3.3

starts with a vowel
```{r}

str_subset(a, "^[aeiou]")

```

starts with a vowel
```{r}

str_subset(a, "^[aeiou]")

```

only contain consonants... stumped on this one, error?
```{r}
#starts with a consonant...
n = 'xcd'
str_match(n, "^[^aeiou].*")

#only has consonants...
n = 'xcd'
head(str_subset(a, "[^aeiou]"))

```
 
 End with ed but not with eed... not using regular expression technically the whole way through but whatever
 
 
```{r}

a[str_detect(a, pattern = 'ed$') & !str_detect(a, pattern = 'eed$')]

```

End with ing or ise 
```{r}
str_subset(a, pattern = 'ing$|ise$')
```

Empirically verify rule that i before e except after c

```{r}

str_subset(a, pattern = 'ie|cei')

```

Is q always followed by u?

```{r}

a[str_detect(a, pattern = 'q') & !str_detect(a, pattern = 'qu')]

```

Create a regular expression that will match telephone numbers as commonly written in your country.

```{r}

number = "(805)-284-6864"

str_subset(number, '^\\(')

```

#11.3.5.1

Interesting, so this is saying find things that go 1221
```{r}

j = "aabbnnz"

str_view(fruit,"(.)(.)\\2\\1", match = T)

str_view(fruit,"(..)\\1", match = T)


```

Start and end with the same character

```{r}
str_subset(c('aba','can'), pattern = "^(.).*\\1$")

```
Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.)

```{r}

str_subset(fruit, pattern = "(..)\\1")

```

Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.)

```{r}
str_subset('eleven',".*(.).*\\1.*\\1")
```

# 11.4 Tools

## 11.4.2

Find all words that start or end with a.

```{r}

str_subset(words, pattern = "^a|a$")

```

Find all words that start with a vowel and end with a consonant.

```{r}

words[str_detect(words,'^[aeiou]') & str_detect(words,'[^aeiou]$')]

```
Are there any words that contain at least one of each different vowel?

```{r}
 word_frame = data_frame(words = words)

word_frame %>% 
  mutate(num_unique_vowels = str_detect(words,'a') + str_detect(words,'e') + 
           str_detect(words,'i') + str_detect(words,'o') + str_detect(words,'u'),
         num_chars = str_count(words, pattern = '.'),
         num_vowels = str_count(words, 'a|e|i|o|u'),
         perc_vowels = num_vowels/num_chars) %>% 
  arrange(desc(num_vowels))
```

# 11.4.3 Extract matches

```{r}
head(sentences)

colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
str_view_all(sentences, colour_match)


has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)

```

# 11.4.3.1

In the previous example, you might have noticed that the regular expression matched “flickered”, which is not a colour. Modify the regex to fix the problem.


`\b` is a boundary, so called in regex by `\\b` ; this allows you to say pick whole words around some thing


```{r}

huh = c('this is red','flickred')

 new_reg <- paste('\\b(',colour_match,')\\b', sep = '')

str_subset(sentences, pattern = new_reg)


```

From the Harvard sentences data, extract:

The first word from each sentence.
All words ending in ing.
All plurals.

Aha! these things are case sensitive, so "a" won't match with "A"

```{r}
head(str_extract(tolower(sentences), "\\b[a-z|']+\\b"))

str_view(sentences, pattern = '\\b[a-z]+ing\\b', match = T)

str_view(sentences, pattern = '\\b[a-z]+[^aeiou]s\\b', match = T) #grab all words that end with an s preceded by a consontant 

```

# 11.4.4

This is interesting, this is something like any word `([^ ]+)`. Meaning, take things that aren't blank, of any length

```{r}

noun <- "(a|the) ([^ ]+)"

has_noun <- sentences %>%
  str_subset(noun) %>%
  head(10)
has_noun %>% 
  str_extract(noun)

tibble(sentence = sentences) %>% 
  tidyr::extract(
    sentence, c("article", "noun"), "(a|the) ([^ ]+)", 
    remove = FALSE
  )

```

## 11.4.4.1 Exercises

Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.

```{r}

numbers <- str_c(c('one','two','three','four','five','six','seven','eight','nine','ten'), collapse = '|')


head(str_match(sentences, pattern = paste('(',numbers,') ([^ ]+)', sep = '')))

head(str_extract(sentences, pattern = paste('(',numbers,') ([^ ]+)', sep = '')))

str_match(sentences, pattern = "(\\b[a-z]+)(')([a-z]+\\b)") #find and separate contractions

```

# 11.3.4 Replacing matches

## 11.4.5.1 Exercises

Replace all forward slashes in a string with backslashes.

Implement a simple version of str_to_lower() using replace_all().

Switch the first and last letters in words. Which of those strings are still words?

```{r}

thing <- c('This is /not ok')

b = writeLines("\\")

str_replace(thing, "/","\\")


lower_rules = letters

names(lower_rules) = LETTERS


lower_fun <- function(wordies) {
  lower_rules = letters
  
  names(lower_rules) = LETTERS
  
  return(str_replace_all(wordies, lower_rules))
  
}

lower_fun('ApPle')


rev_words <-  (str_replace(tolower(words), "(^[a-z])([a-z]+)([a-z]$)",'\\3\\2\\1')) #Switch the first and last letters in words. Which of those strings are still words?

rev_words[rev_words %in% words]            
            

```

# 11.4.6 Splitting

```{r}

sentences %>%
  head(5) %>% 
  str_split(" ")

"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]


sentences %>%
  head(5) %>% 
  str_split(" ", simplify = T)

fields <- c("Name: Hadley", "Country: NZ", "Age: 35")
fields %>% str_split(": ", n = 2, simplify = TRUE)
```

```{r}
x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))

str_split(x, " ")[[1]]

str_split(x, boundary("word"))[[1]]
```



```{python}

a = {'New York': 'USA','Mexico City':'Mexico'}

print(a['Mexico City'])

```

## 11.4.6.1 String splitting exercises

Split up a string like "apples, pears, and bananas" into individual components.

Why is it better to split up by boundary("word") than " "? **since it catches things like ","


What does splitting with an empty string ("") do? Experiment, and then read the documentation.

```{r}

a <- c('apples, pears, and bananas')

str_split(a, ' ')[[1]]

str_split(a, boundary('word'))[[1]]

str_split(a, '')[[1]]


```

## 14.5.1 Exercises

How would you find all strings containing \ with regex() vs. with fixed()?

What are the five most common words in sentences?

```{r}

test = c('Hell\\o','hello','te\\st')

str_subset(test,'\\\\')

str_detect(test, fixed("\\"))


#most common words in sentances

split_words <- str_split(sentences, boundary('word'), simplify = T) %>% 
  as_data_frame() %>% 
  gather('position','word') %>% 
  mutate(word = tolower(word)) %>% 
  count(word) %>% 
  arrange(desc(n))
```

# Factors


```{r}

library(forcats)

forcats::gss_cat

ggplot(gss_cat, aes(race)) + 
  geom_bar() + 
  scale_x_discrete(drop = FALSE)


relig <- gss_cat %>% 
  group_by(relig) %>% 
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )

ggplot(relig, aes(tvhours, relig)) + geom_point()
ggplot(relig, aes(tvhours, fct_reorder(relig, tvhours))) + geom_point()
```

```{r}

gss_cat$marital

fct_relevel(gss_cat$marital, 'Divorced') #if you just want to pull a few things forward and leave the rest alone


by_year <- gss_cat %>% 
  group_by(year, marital) %>% 
  count() %>% 
  mutate(prop = n / sum(n))

ggplot(by_year, aes(year, prop, colour = marital)) + 
  geom_line()

ggplot(by_year, aes(year, prop, colour = fct_reorder2(marital, year, prop))) + 
  geom_line()

```

